---
title: "Bivariate Analysis For Banking Data"
author: "Ada Lazuli"
date: '2022-09-05'
output:
  html_document:
    toc: yes
    toc_depth: 1
    code_folding: show
  pdf_document:
    toc: yes
    toc_depth: '1'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r, out.width = '100%', warning=FALSE, message=FALSE}
library(ggplot2)
library(ggcorrplot)
library(dplyr)
library(tidyverse)
```

# Helper Functions
```{r, out.width = '100%', warning=FALSE, message=FALSE}
box_cat_plt <- function(df, x, y, xname, yname) {
  title <- paste("Boxplot Comparing", xname, "to", yname)
  ggplot(df, aes(x = x, color = y,  fill = y)) + geom_boxplot(alpha = 0.5) + coord_flip() +
    xlab(xname) +
    theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
          axis.title.y = element_blank()) + 
    ggtitle(title) + 
    labs(fill = yname, color = yname)
}

scatter_comp_plt <- function(df, x, y, color, colorname){
  title <- paste("Scatter Plot of AGE by Limit with", colorname, "shading")
  ggplot(df, aes(x = x, y = y, color = color)) +
    geom_point(alpha = 0.3) +
    xlab("Age") +
    ylab("Credit Limit") +
    ggtitle(title)
}

chi_matrix <- function(df) {
  rtn <- c()
  for(x in colnames(df)){
    temp <- c()
    for (y in colnames(df)) {
      temp <- c(temp, chisq.test(df[[x]], df[[y]])$p.value)
    }
    rtn <- rbind(rtn, temp)
  }
  
  rtn <- as.data.frame(rtn)
  colnames(rtn) <- colnames(df)
  rownames(rtn) <- colnames(df)
  
  return(rtn)
}

chi_vector <- function(df, target) {
  rtn <- c()
  status <- c()
  for(x in colnames(df)){
    val <-  round(chisq.test(df[[x]], target)$p.value,digits = 4)
    rtn <- c(rtn, val)
    status <- c(status, if(val < 0.01) "Passed" else "Failed")
  }
  rtn <- cbind(rtn,status)
  rtn <- cbind(rtn, colnames(df))
  rtn <- as.data.frame(rtn)
  colnames(rtn) <- c("pvalue", "Result", "feature")
  return(rtn)
}

extract_month <- function(x) {
  if (x == "PAY_0"){
    return("9")
  } else if (x == "PAY_2") {
    return("8")
  }else if (x == "PAY_3") {
    return("7")
  }else if (x == "PAY_4") {
    return("6")
  }else if (x == "PAY_5") {
    return("5")
  }else if (x == "PAY_6") {
    return("4")
  } else if (x == "BILL_AMT1") {
    return("9")
  } else if (x == "BILL_AMT2") {
    return("8")
  }else if (x == "BILL_AMT3") {
    return("7")
  }else if (x == "BILL_AMT4") {
    return("6")
  }else if (x == "BILL_AMT5") {
    return("5")
  }else if (x == "BILL_AMT6") {
    return("4")
  }else if (x == "PAY_AMT1") {
    return("9")
  }else if (x == "PAY_AMT2") {
    return("8")
  }else if (x == "PAY_AMT3") {
    return("7")
  }else if (x == "PAY_AMT4") {
    return("6")
  }else if (x == "PAY_AMT5") {
    return("5")
  }else if (x == "PAY_AMT6") {
    return("4")
  }
  
}

```

# Data Loading

```{r, out.width = '100%', warning=FALSE, message=FALSE}
df <- read.csv("credit card default data set.csv")
df$default <- factor(df$default.payment.next.month, labels = c("No", "Yes"))
df$default.payment.next.month <- NULL
```

# Pre-processing

The following features are categorical and need to be converted into factor types:

1. SEX
2. EDUCATION
3. MARRIAGE

```{r, out.width = '100%', warning=FALSE, message=FALSE}
df$SEX <- factor(df$SEX, labels = c("Male", "Female"))
df$EDUCATION <- factor(df$EDUCATION, labels = c("Others", "Graduate School", "Univserity", "High School", "Others",  "Others",  "Others"))
df$MARRIAGE <- factor(df$MARRIAGE, labels = c("Others", "Married", "Single","Others"))
```

# Bivariate Analysis (Non-Payment Features)

## Overview

The dataset has __`r nrow(df)`__ rows and __`r ncol(df)`__ features. The names of the features are:

_`r colnames(df)`_

```{r, out.width = '100%', warning=FALSE, message=FALSE}
glimpse(df)
```

## LIMIT_BAL v.s. Categorical Features

The P-value for chi-squared test for independence between LIMIT_BAL and each category is:

1. LIMIT_BAL vs. SEX: `r chisq.test(df$LIMIT_BAL, df$SEX)$p.value`
1. LIMIT_BAL vs. EDUCATION: `r chisq.test(df$LIMIT_BAL, df$EDUCATION)$p.value`
1. LIMIT_BAL vs. MARRIAGE: `r chisq.test(df$LIMIT_BAL, df$MARRIAGE)$p.value`
1. LIMIT_BAL vs. Default: `r chisq.test(df$LIMIT_BAL, df$default)$p.value`



```{r, out.width = '100%', warning=FALSE, message=FALSE}
box_cat_plt(df, df$LIMIT_BAL, df$SEX, "Credit Limit", "SEX")
box_cat_plt(df, df$LIMIT_BAL, df$EDUCATION, "Credit Limit", "EDUCATION")
box_cat_plt(df, df$LIMIT_BAL, df$MARRIAGE, "Credit Limit", "MARRIAGE")
```

## AGE v.s. Categorical Features

The P-value for chi-squared test for independence between AGE and each category is:

1. AGE vs. SEX: `r chisq.test(df$AGE, df$SEX)$p.value`
1. AGE vs. EDUCATION: `r chisq.test(df$AGE, df$EDUCATION)$p.value`
1. AGE vs. MARRIAGE: `r chisq.test(df$AGE, df$MARRIAGE)$p.value`
1. AGE vs. Default: `r chisq.test(df$AGE, df$default)$p.value`

```{r, out.width = '100%', warning=FALSE, message=FALSE}
box_cat_plt(df, df$AGE, df$SEX, "AGE", "SEX")
box_cat_plt(df, df$AGE, df$EDUCATION, "AGE", "EDUCATION")
box_cat_plt(df, df$AGE, df$MARRIAGE, "AGE", "MARRIAGE")
```

## AGE v.s. LIMIT_BAL

The correlation between AGE and LIMIT_BAL is `r cor(df$AGE, df$LIMIT_BAL)`

```{r, out.width = '100%', warning=FALSE, message=FALSE}
scatter_comp_plt(df, df$AGE, df$LIMIT_BAL, df$SEX, "SEX")
scatter_comp_plt(df, df$AGE, df$LIMIT_BAL, df$EDUCATION, "EDUCATION")
scatter_comp_plt(df, df$AGE, df$LIMIT_BAL, df$MARRIAGE, "MARRIAGE")
scatter_comp_plt(df, df$AGE, df$LIMIT_BAL, df$default, "Defaults")
```

## Categorical Chi Squre Matrix

A matrix of the results for chi-squared testing was created to evalute the categorical features against each other for independence.
```{r, out.width = '100%', warning=FALSE, message=FALSE}
chi_matrix(df %>% select(SEX, MARRIAGE, EDUCATION, default))
```


# Bivariate Analysis (Payment Features)

## Correlation Matrix
```{r, out.width = '100%', warning=FALSE, message=FALSE}
ggcorrplot(cor(df%>% select(-AGE, -LIMIT_BAL, -SEX, -EDUCATION, -MARRIAGE, -default, -ID)))
```

## Independence Testing

```{r, out.width = '100%', warning=FALSE, message=FALSE}
chiv <- chi_vector(df%>% select(-AGE, -LIMIT_BAL, -SEX, -EDUCATION, -MARRIAGE, -default, -ID), df$default)
ggplot(chiv, aes(x = feature, y = pvalue, color = Result, fill = Result)) + geom_col() +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = -0.05))
```


# Bivariate Analysis (Time Series)

## Pre-Processing

```{r, out.width = '100%', warning=FALSE, message=FALSE}

df$Index <- 1:nrow(df)

df_hist <- df %>% pivot_longer(c(PAY_0, PAY_2, PAY_3, PAY_4, PAY_5, PAY_6), "Month") %>%
  rename(HIST = value)

df_hist$Month <- sapply(df_hist$Month, extract_month)

df_billed <- df %>%   pivot_longer(c(BILL_AMT1, BILL_AMT2, BILL_AMT3, BILL_AMT4, BILL_AMT5, BILL_AMT6), "Month") %>%
  rename(BILLED = value) %>%
  select(Month, BILLED, Index)

df_billed$Month <- sapply(df_billed$Month, extract_month)

df_paid <- df %>% pivot_longer(c(PAY_AMT1, PAY_AMT2, PAY_AMT3, PAY_AMT4, PAY_AMT5, PAY_AMT6), "Month") %>%
  rename(PAID = value) %>%
  select(Month, PAID, Index)

df_paid$Month <- sapply(df_paid$Month, extract_month)

df_joined <- inner_join(x = df_billed, y = df_paid, by = c("Index", "Month")) %>% 
inner_join(df_hist, by = c("Index", "Month")) %>%
select(Month, BILLED, PAID, HIST, LIMIT_BAL, SEX, EDUCATION, MARRIAGE, AGE, default)

df_sample <- df_joined %>% group_by(default) %>% slice_sample(prop = .3) %>% ungroup()
```

## Quantitivative Feature Chi Square Eval

```{r, out.width = '100%', warning=FALSE, message=FALSE}
chiv <- chi_vector(df_sample%>% select(-SEX, -EDUCATION, -MARRIAGE, -default), df_sample$default)
ggplot(chiv, aes(x = feature, y = pvalue, color = Result, fill = Result)) + geom_col() +
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust = -0.05))
```

## Qualitative Variable Correlation

```{r, out.width = '100%', warning=FALSE, message=FALSE}
ggcorrplot(cor(df_sample %>% select(AGE, LIMIT_BAL, HIST, BILLED, PAID)))
```

## HIST over Time

```{r, out.width = '100%', warning=FALSE, message=FALSE}
ggplot(df_sample, aes(x = Month, y = HIST, color = default)) + geom_point() + geom_jitter(alpha = .02)
```

## BILLED over Time
```{r, out.width = '100%', warning=FALSE, message=FALSE}
ggplot(df_sample, aes(x = Month, y = BILLED, color = default)) + geom_point() + geom_jitter(alpha = .02)
```

## Paid over Time
```{r, out.width = '100%', warning=FALSE, message=FALSE}
 ggplot(df_sample, aes(x = Month, y = PAID, color = default)) + geom_point() + geom_jitter(alpha = .02)
```

 ## PAID vs BILLED
```{r, out.width = '100%', warning=FALSE, message=FALSE}
 ggplot(df_sample, aes(x = BILLED, y = PAID, color = default)) + geom_point()

 ggplot(df_sample, aes(x = BILLED, y = PAID, color = default)) + geom_point() + scale_y_log10() + scale_x_log10()
```


