---
title: "Univariate Analysis For Banking Data"
author: "Ada Lazuli"
date: '2022-09-05'
output:
  pdf_document:
    toc: yes
    toc_depth: '1'
  html_document:
    toc: yes
    toc_depth: 1
    code_folding: show
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r, out.width = '100%', warning=FALSE, message=FALSE}
library(e1071)
library(ggplot2)
library(dplyr)
```

# Helper Functions

```{r, out.width = '100%', warning=FALSE, message=FALSE}
base_hist <- function(df, x, target_name){
  title <- paste("Histogram of ", target_name)
  ggplot(df, aes(x = x)) + geom_histogram(alpha = 0.4) +
    ggtitle(title) +
    xlab(target_name) +
    ylab("Frequency Count")
}

zero_plot <- function(df, x, target_name) {
  title <- paste("Zero Plot of ", target_name)
  ggplot(df, aes(x = x, y = 0, color =  default.payment.next.month)) + 
    geom_point(alpha = 0.45) +
    ggtitle(title) +
    xlab(target_name) +
    theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
          axis.title.y = element_blank()) + 
    labs(color = "Defaults")
}

boxplot_comp <- function(df, x, target_name) {
  title <- paste("Boxplot of ", target_name)
  ggplot(df, aes(x = x, fill = default.payment.next.month)) + 
    geom_boxplot(alpha = 0.45) +
    ggtitle(title) +
    xlab(target_name) +
    theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
          axis.title.y = element_blank()) + 
    labs(fill = "Defaults")
}

bar_comp_plot <- function(df, x, target_name) {
  title <- paste("Barplot of ", target_name)
  ggplot(df, aes(x = x, fill =  default.payment.next.month, color = default.payment.next.month)) + 
    geom_bar(alpha = 0.5, position = "dodge") +
    ggtitle(title) +
    xlab(target_name) +
    ylab("Frequency Count") +
    labs(fill = "Defaults", color = "Defaults")
}
```


# Data Loading

```{r, out.width = '100%', warning=FALSE, message=FALSE}
df <- read.csv("credit card default data set.csv")
df$default.payment.next.month <- factor(df$default.payment.next.month, labels = c("No", "Yes"))
```

# Univariate Analysis

## Overview

The dataset has __`r nrow(df)`__ rows and __`r ncol(df)`__ features. The names of the features are:

_`r colnames(df)`_

```{r, out.width = '100%', warning=FALSE, message=FALSE}
glimpse(df)
```


## ID

The ID feature appears to be an index with a minimum of __`r min(df$ID)`__ and maximum of __`r max(df$ID)`__.

## LIMIT_BAL (X1)

X1 is the first feature for predictive modeling and consists for the amount of credit given that accounts for both the individual credit and family credit.

The variable appears to consist of integers, with a maximum of __`r max(df$LIMIT_BAL)`__ and a minimum of __`r min(df$LIMIT_BAL)`__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$LIMIT_BAL)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$LIMIT_BAL
name <- "Credit Limit (X1)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```

## SEX (X2)

The gender variable refers to the gender of the individual, with a 1 for male and a 2 for female. 

```{r, out.width = '100%', warning=FALSE, message=FALSE}
df$SEX <- factor(df$SEX, labels = c("Male", "Female"))
table(df$SEX)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$SEX
name <- "SEX (X2)"
bar_comp_plot(df,x,name)
```

## EDUCATION (X3)

The education feature captures the education level of the individual. Possible values are:
1. Graduate School
2. University
3. High School
4. Others


__NOTE:__ Values of 0, 5, and 6 were found in the data and added to the _Others_ category.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
df$EDUCATION <- factor(df$EDUCATION, labels = c("Others", "Graduate School", "Univserity", "High School", "Others",  "Others",  "Others"))
table(df$EDUCATION)
```

### Graphical Analysis

```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$EDUCATION
name <- "EDUCATION (X3)"
bar_comp_plot(df,x,name)
```

## MARRIAGE (X4)

The marriage feature captures the marital status of the individual. Possible values are:
1. Married
2. Single
3. Others

```{r, out.width = '100%', warning=FALSE, message=FALSE}
df$MARRIAGE <- factor(df$MARRIAGE, labels = c("Others", "Married", "Single","Others"))
table(df$MARRIAGE)
```

### Graphical Analysis

```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$MARRIAGE
name <- "MARRIAGE (X4)"
bar_comp_plot(df,x,name)
```

## AGE (X5)

The age feature captures the age of the individual.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$AGE)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$AGE
name <- "AGE (X5)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## PAY_0 (X6)

The PAY_0 feature captures the payment status  of the individual for the month of __September, 2005__. A negative number represents a duly payment, a 0 represents the month not being paid, and a positive number represents the number of months behind payment.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_0)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_0
name <- "PAY_0 (X6)"
bar_comp_plot(df,x,name)
boxplot_comp(df,x,name)
```

## PAY_2 (X7)

The PAY_1 feature captures the payment status  of the individual for the month of __August, 2005__. A negative number represents a duly payment, a 0 represents the month not being paid, and a positive number represents the number of months behind payment.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_2)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_2
name <- "PAY_2 (X7)"
bar_comp_plot(df,x,name)
boxplot_comp(df,x,name)
```

## PAY_3 (X8)

The PAY_3 feature captures the payment status  of the individual for the month of __July, 2005__. A negative number represents a duly payment, a 0 represents the month not being paid, and a positive number represents the number of months behind payment.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_3)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_3
name <- "PAY_3 (X8)"
bar_comp_plot(df,x,name)
boxplot_comp(df,x,name)
```

## PAY_4 (X9)

The PAY_4 feature captures the payment status  of the individual for the month of __June, 2005__. A negative number represents a duly payment, a 0 represents the month not being paid, and a positive number represents the number of months behind payment.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_4)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_4
name <- "PAY_4 (X9)"
bar_comp_plot(df,x,name)
boxplot_comp(df,x,name)
```

## PAY_5 (X10)

The PAY_5 feature captures the payment status  of the individual for the month of __May, 2005__. A negative number represents a duly payment, a 0 represents the month not being paid, and a positive number represents the number of months behind payment.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_5)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_5
name <- "PAY_5 (X10)"
bar_comp_plot(df,x,name)
boxplot_comp(df,x,name)
```

## PAY_6 (X11)

The PAY_6 feature captures the payment status  of the individual for the month of __April, 2005__. A negative number represents a duly payment, a 0 represents the month not being paid, and a positive number represents the number of months behind payment.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_6)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_6
name <- "PAY_6 (X11)"
bar_comp_plot(df,x,name)
boxplot_comp(df,x,name)
```


## BILL_AMT1 (X12)

The BILL_AMT1 feature captures the billed amount to the individual for the month of __September, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$BILL_AMT1)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$BILL_AMT1
name <- "BILL_AMT1 (X12)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```

## BILL_AMT2 (X13)

The BILL_AMT2 feature captures the billed amount to the individual for the month of __August, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$BILL_AMT2)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$BILL_AMT2
name <- "BILL_AMT2 (X13)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## BILL_AMT3 (X14)

The BILL_AMT3 feature captures the billed amount to the individual for the month of __July, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$BILL_AMT3)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$BILL_AMT3
name <- "BILL_AMT3 (X14)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## BILL_AMT4 (X15)

The BILL_AMT4 feature captures the billed amount to the individual for the month of __June, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$BILL_AMT4)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$BILL_AMT4
name <- "BILL_AMT4 (X15)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```

## BILL_AMT5 (X16)

The BILL_AMT5 feature captures the billed amount to the individual for the month of __May, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$BILL_AMT5)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$BILL_AMT5
name <- "BILL_AMT5 (X16)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```

## BILL_AMT6 (X17)

The BILL_AMT6 feature captures the billed amount to the individual for the month of __April, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$BILL_AMT6)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$BILL_AMT6
name <- "BILL_AMT6 (X17)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## PAY_AMT1 (X18)

The PAY_AMT1 feature captures the payment amount from the individual for the month of __September, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_AMT1)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_AMT1
name <- "PAY_AMT1 (X18)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## PAY_AMT2 (X19)

The PAY_AMT2 feature captures the payment amount from the individual for the month of __August, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_AMT2)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_AMT2
name <- "PAY_AMT2 (X19)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```



## PAY_AMT3 (X20)

The PAY_AMT3 feature captures the payment amount from the individual for the month of __July, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_AMT3)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_AMT3
name <- "PAY_AMT3 (X20)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## PAY_AMT4 (X21)

The PAY_AMT4 feature captures the payment amount from the individual for the month of __June, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_AMT4)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_AMT4
name <- "PAY_AMT4 (X21)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```




## PAY_AMT5 (X22)

The PAY_AMT5 feature captures the payment amount from the individual for the month of __May, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_AMT5)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_AMT5
name <- "PAY_AMT5 (X22)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## PAY_AMT6 (X23)

The PAY_AMT6 feature captures the payment amount from the individual for the month of __April, 2005__.

```{r, out.width = '100%', warning=FALSE, message=FALSE}
summary(df$PAY_AMT6)
```

### Graphical Analysis
```{r, out.width = '100%', warning=FALSE, message=FALSE}
x <- df$PAY_AMT6
name <- "PAY_AMT6 (X23)"
base_hist(df,x,name)
zero_plot(df,x,name)
boxplot_comp(df,x,name)
```


## Defaults (Y1)

The target variable for the dataset is whether the individual defaults on a payment. 

```{r, out.width = '100%', warning=FALSE, message=FALSE}
table(df$default.payment.next.month)
```












